---
title: "Software tutorials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Software tutorials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial, we demonstrate basic use of the SpikeInference package.

First load the pacakge:
```{r}
require(SpikeInference)
```

To illustrate the use, we will first simulate from a AR-1 model for calcium imaging data according to 
$$Y_t = c_t + \epsilon_t, \epsilon_t \sim N(0, \sigma^2),$$ where $$c_t = \gamma  c_{t-1} + z_t, z_t\geq 0$$. We will use $\sigma = 0.1, \gamma = 0.98, z_t \overset{i.i.d.}{\sim}\text{Poisson}(0.01)$ for this particular example.

```{r}
gam <- 0.98
sigma <- 0.3
n_length <- 1000
curr_sim <- simulate_ar1(n = n_length, gam = gam, poisMean = 0.01, sd = sigma, seed = 2)
plot(curr_sim)
```



#### $\ell_0$ spike estimation 
We estimate spikes using an $\ell_0$ penalty with tuning parameter $\lambda=4$:
```{r}
LAMBDA <- 0.7
fit_spike <- spike_estimates(dat = curr_sim$fl, decay_rate = gam, 
                             tuning_parameter = LAMBDA, 
                             functional_pruning_out = FALSE)
plot(fit_spike)
```

The `fit_spike` object contains spike fit information:
```{r}
str(fit_spike)
```

Importantly, the first five estimated spike locations are: 
```{r}
(fit_spike$spikes)[1:5]
```

#### $\ell_0$ spike inference 

In this section we demonstrate how to use our software to obtain p-values and confidence intervals for estimated spikes with 2 different window size h (1 and 10); for displaying purposes, we will only consider the first five estimated spikes.

Recall that the p-values we are interested in have the form
$$\mathbb{P}\left(\phi \geq \nu^\top y \middle | \phi \in \mathcal{S}\cap(0,\infty)\right)$$, for a given $\nu$.

#### For h = 1:
```{r}
h <- 1
inference_spike_toy_example_h_1 <- spike_inference(dat = curr_sim$fl, decay_rate = gam,
                                       tuning_parameter = LAMBDA, window_size = h, 
                                       sig = sigma*sigma, return_conditioning_sets = FALSE,return_ci = TRUE,lower_trunc=0)

knitr::kable(data.frame(estimated_spikes = inference_spike_toy_example_h_1$spikes[1:5], 
                        pvals = inference_spike_toy_example_h_1$pvals[1:5],
                        LCB = inference_spike_toy_example_h_1$LCB[1:5],
                        UCB = inference_spike_toy_example_h_1$UCB[1:5]))

```

#### For h = 10:
```{r}
h <- 10
inference_spike_toy_example_h_10 <- spike_inference(dat = curr_sim$fl, decay_rate = gam,
                                       tuning_parameter = LAMBDA, window_size = h, 
                                       sig = sigma*sigma, return_conditioning_sets = FALSE,return_ci = TRUE,lower_trunc=0)

knitr::kable(data.frame(estimated_spikes = inference_spike_toy_example_h_10$spikes[1:5], 
                        pvals = inference_spike_toy_example_h_10$pvals[1:5],
                        LCB = inference_spike_toy_example_h_10$LCB[1:5],
                        UCB = inference_spike_toy_example_h_10$UCB[1:5]))
```

We see the drastic change in p-values (power) for a larger h!



